The semantic drift of quotations in blogspace: a case study in short-term cultural evolution

### 2016-07-13 brains-protocol.tex
style/protocol: missing \newtext's and left-over self-notes

Manual inspection of this data set revealed that it contains a significant number of everyday language quotations (such as ``it was much better than I expected'', ``did that just happen'', as well as many simple expletive-based sentences).
Their presence is largely due to random variations around casual expressions, while we are interested in transformations of news-related quotes causally linked to an original, identifiable utterance.
To filter them out, we exclude quotes with less than 5 words or \newtext{whose occurrences span more} than 80 days (indicating causally unrelated occurrences), as well as quotes not written in English.
\newtext{Clusters that are emptied by this procedure are therefore excluded.}
If, after this screening, a cluster's occurrences still span more than 80 days (because of short-lived but unrelated quotes far apart in time), we also exclude it.
\newtext{We eventually keep 50,427 clusters (out of 71,568; \hbox{i.e.} 70.5\%), containing a total of 141,324 unique quotes (out of 310,457; \hbox{i.e.} 45.5\%) making up about 2.60m occurrences (out of 7.67m; \hbox{i.e.} 33.9\%).\footnote{
The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is to be expected for everyday language utterances.
}
Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (one such lost quote, for instance, is "the city is tired of me and the organization and I have run our course together"), we check that our approach fulfills its goals by coding a random sub-sample of 100 clusters:
35 of them are rejected by the filter, with 15 false negatives (rejected clusters that should have been kept) and 9 false positives (clusters kept when they should have been rejected), giving a precision score of 0.862 and a recall score of 0.789.
Furthermore, all but one of the 9 false positives are left with a single non-rejected quote, meaning those clusters are ignored by our substitution analysis; this brings the effective precision of our filter to 0.982.\footnote{
A similar analysis was made for language detection, which is part of the cluster filtering:
out of 100 randomly sampled quotes, 17 are rejected because their detected language is not English, with no false positives and 6 false negatives, giving a precision score of 1 and a recall score of 0.933.
Of the 6 false negatives, 4 had less than 5 tokens and would have been excluded by the cluster filter anyway.
}}

@@

\newtext{Several of these features are strongly related and can be grouped together.
To make correlation values as well as future comparisons more reliable, we log-transformed features that have marked exponential distributions (a few words valued orders of magnitude higher than the vast majority of other words).\footnote{
The distributions of original and log-transformed features appears in the supplementary material, along with correlation scatter plots for each feature pair, summarized below with correlation values.
}}

@@

Let us give an example:
\newtext{say the quotation "These accusations are false and \textbf{absurd}" ($q$) appears in two different blogs on January 19, and the slightly different quotation "These accusations are false and \textbf{incoherent}" ($q'$) appears in another blog on the 20th of January.
The second occurrence of $q$ can safely be assumed to be a faithful copy of the first one the same day.
And since $q$ is fairly prominent when $q'$ first appears, we could assume that the author of $q'$ on the 20th based herself on $q$ as is shown with a dashed line in Fig.~\ref{fig:substitution-unmodelled}.
Now say a third version, "These \textbf{allegations} are false and \textbf{incoherent}" ($q''$) also appears once on January 19 and once on January 20 after $q'$.
$q$ and $q''$ differ by two substitutions, so we ignore the possibility that one was written based on the other.\footnote{
NOTE WE TESTED THIS.
}
$q''$ is only one substitution away from $q'$ however, so we could also consider the first occurrence of $q''$ as a potential source for $q'$ on the 20th.
Conversely, the occurrence of $q''$ on the 20th could be considered as a substitution from $q'$, or as a faithful copy from its initial occurrence on January 19.}
(Options shown in Fig.~\ref{fig:substitution-unmodelled}.)

>>>>

Manual inspection of this data set revealed that it contains a significant number of everyday language quotations (such as ``it was much better than I expected'', ``did that just happen'', as well as many simple expletive-based sentences).
Their presence is largely due to random variations around casual expressions, while we are interested in transformations of news-related quotes causally linked to an original, identifiable utterance.
To filter them out, we exclude quotes with less than 5 words or \newtext{whose occurrences span more} than 80 days (indicating causally unrelated occurrences), as well as quotes not written in English.
\newtext{Clusters that are emptied by this procedure are therefore excluded.}
If, after this screening, a cluster's occurrences still span more than 80 days (because of short-lived but unrelated quotes far apart in time), we also exclude it.
\newtext{We eventually keep 50,427 clusters (out of 71,568; \hbox{i.e.} 70.5\%), containing a total of 141,324 unique quotes (out of 310,457; \hbox{i.e.} 45.5\%) making up about 2.60m occurrences (out of 7.67m; \hbox{i.e.} 33.9\%).\footnote{
The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is to be expected for everyday language utterances.
}
Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (one such lost quote, for instance, is "the city is tired of me and the organization and I have run our course together"), we check that our approach fulfills its goals by coding a random sub-sample of 100 clusters:
35 of them are rejected by the filter, with 15 false negatives (rejected clusters that should have been kept) and 9 false positives (clusters kept when they should have been rejected), giving a precision score of 0.862 and a recall score of 0.789.
Furthermore, all but one of the 9 false positives are left with a single non-rejected quote, meaning those clusters are ignored by our substitution analysis; this brings the effective precision of our filter to 0.982.\footnote{
\newtext{A similar analysis was made for language detection, which is part of the cluster filtering:
out of 100 randomly sampled quotes, 17 are rejected because their detected language is not English, with no false positives and 6 false negatives, giving a precision score of 1 and a recall score of 0.933.
Of the 6 false negatives, 4 had less than 5 tokens and would have been excluded by the cluster filter anyway.
}}}

@@

\newtext{Several of these features are strongly related and can be grouped together.
To make correlation values as well as future comparisons more reliable, we log-transformed features that have marked exponential distributions (a few words valued orders of magnitude higher than the vast majority of other words).\footnote{
\newtext{The distributions of original and log-transformed features appears in the supplementary material, along with scatter plots for each feature pair, summarized below with correlation values.
}}}

@@

Let us give an example:
\newtext{say the quotation "These accusations are false and \textbf{absurd}" ($q$) appears in two different blogs on January 19, and the slightly different quotation "These accusations are false and \textbf{incoherent}" ($q'$) appears in another blog on the 20th of January.
The second occurrence of $q$ can safely be assumed to be a faithful copy of the first one the same day.
And since $q$ is fairly prominent when $q'$ first appears, we could assume that the author of $q'$ on the 20th based herself on $q$ as is shown with a dashed line in Fig.~\ref{fig:substitution-unmodelled}.
Now say a third version, "These \textbf{allegations} are false and \textbf{incoherent}" ($q''$) also appears once on January 19 and once on January 20 after $q'$.
$q$ and $q''$ differ by two substitutions, so we discard the possibility that one was written based on the other (this below for further details).
$q''$ is only one substitution away from $q'$ however, so we could also consider the first occurrence of $q''$ as a potential source for $q'$ on the 20th.
Conversely, the occurrence of $q''$ on the 20th could be considered as a substitution from $q'$, or as a faithful copy from its initial occurrence on January 19.}
(Options shown in Fig.~\ref{fig:substitution-unmodelled}.)
